Optimal strategies of influence in structured social groups

### 2015-07-17 paper/manuscript.tex
updated reports and paper
updated paper

\begin{document}
%\begin{frontmatter}
%\preprint{}
\title{Optimal strategies of influence in structured populations}
\author{Colin J. Torney}
\affiliation{Centre for Mathematics \& the Environment, University of Exeter, Penryn Campus, Cornwall, UK}
%\affiliation{Department of Ecology \& Evolutionary Biology, Princeton University, \\Princeton, New Jersey 08544}
\author{Simon A. Levin}
\affiliation{Department of Ecology \& Evolutionary Biology, Princeton University, \\Princeton, New Jersey 08544}

@@

Each individual, $i$, holds an opinion $S_i \in \{ -1,1\}$ that is observable by others in its local neighborhood and is correct when $S_i = \mbox{sign}(G)$. Individuals update their states based on social and personal information and a Bayesian decision rule \cite{nitzan1982optimal,perez2011collective, Perreault, torney2015} is employed. This means that an individual will switch to the state that is the most likely to be correct, given all available information. Formally, individual $i$ will adopt the state $S_i=1$ if
\begin{equation}
 P\left(G \ge 0\big| g_i, \mathbf{S}_{j\in \mathcal{N}_i}\right)  > P\left(G < 0\big| g_i, \mathbf{S}_{j\in \mathcal{N}_i}\right)
\label{opt_sol1}
\end{equation}
where $\mathcal{N}_i$ defines the social interaction neighborhood. To impose a decision rule as a function of $g_i$ we rearrange Eqn.~\ref{opt_sol1} using Bayes' rule, assume individuals place equal weighting on the opinions of neighbors, that they know the level of noise in the environmental cue, and that it is symmetric around zero. Eqn.~\ref{opt_sol1} may then be written as,
\begin{equation}
\frac{\exp{ \left( - \frac{\vert g_i -  \infty \vert}{\sigma} \right)}}{\exp{ \left( - \frac{\vert g_i + \infty \vert}{\sigma} \right)}} > \chi^{1-2x_L}.
\label{rule}
\end{equation}
Here we have denoted the fraction of neighbors an individual observes in the state $S=1$ as $x_L$, used $\pm \infty$ as estimates of the actual value of $G$ (these will cancel), and introduced a social weighting variable $\chi$ which may be interpreted as the belief that a single neighbor is correct rescaled by the neighborhood size. This rescaling means decisions are based on the fraction of neighbors that are correct in the local neighborhood and not the absolute number, ensuring a comparison can be made between social networks with varying average degree.

Rearranging Eqn.~\ref{rule} leads to a decision rule such that individual $i$ will adopt the opinion $S_i = 1$ iff
\begin{equation}
g_i >  \sigma\ln\chi  \left(\frac{1}{2}-x_L\right) 
\label{opt_rule}
\end{equation}

@@

Numerical calculations of $x_2-x_1$ and the optimal strategy are shown in Fig.~\ref{figx2x1}. More intuition into the underlying processes that control the distance between the inflexion points may be attained by performing an expansion of the social information, $\mathcal{F}(x)$, in terms of the moments of the sampling process. Taylor expanding the inner term %$\exp{\left[ - \ln\chi  \left(\frac{1}{2}-x_L\right)\right]}$ 
of Eqn.~\ref{social} around $x_L=x$ leads to % allows us to replace the summation over all values of $x_L$ with its mean and variance, i.e.
\begin{equation}
\mathcal{F}(x) = \frac{1}{2}\exp{\left[ - \ln\chi  \left(\frac{1}{2}-x\right)\right]} \left( 1 + \frac{ \ln(\chi)^2}{2}\mbox{Var}(x_L) \right)
\end{equation}
Effectively this equation approximates the process of observing a restricted subset of the population. The observations of opinions of neighbors represents an estimate, $x_L$, of the true population average opinion, $x$. This random process of observations may be approximated in terms of the mean and variance of the sampling. While the underlying network structure is complex and non-Markovian, it may be modelled as correlated binomial, in essence we model the process as sampling with pairwise correlation, when there is transitivity near-neighbors are likely to have copied one another hence opinions will be clustered. Following ~\cite{witt2004moody}, the variance of a correlated binomial is
\begin{equation}
\mbox{Var}(x_L) = x(1-x)\left( \rho + \frac{1-\rho}{N_s} \right)
\end{equation}
where $\rho$ is the pairwise correlation coefficient. Assuming the network structure and conformity create a positive correlation of opinions amongst neighbors, then we may define an increased likelihood that neighbor $j$ is in state $S_j=1$, given $S_k=1$, so that
\begin{equation}
P(S_j=1|S_k=1) = x(1+\alpha)
\end{equation}
where $\alpha$ is determined by the transitivity (probability $k$ is in the neighborhood of $j$) and conformity. This means 
\begin{equation}
\rho = \frac{x\alpha}{1-x}
\end{equation}
and
\begin{align}
\mathcal{F}(x) &= \frac{1}{2}\exp{\left[ -  \ln\chi  \left(\frac{1}{2}-x\right)\right]} \times \nonumber \\ & \left( 1 + \frac{ \ln(\chi)^2}{2}\left(\frac{x(1-x(1+\alpha))}{N_s} +  \alpha x^2  \right)   \right).
\end{align}
Solutions to $\mathcal{F}(x)=x$ may be constructed with a perturbation method, with small parameters $\alpha$ and $1/N_s$. We begin with 
\begin{equation}
x_1 \sim x^0_1 + \alpha x^1_1 + 1/N_s x^2_1
\end{equation}
The zeroth order solution is for the mean field where there is no clustering and no noise in the sampling process, hence
\begin{align}
x^0  = \frac{1}{2}\exp{\left[ 1/\sigma -  \ln\chi  \left(\frac{1}{2}-x^0\right)\right]} 
\end{align}
For small x
\begin{equation}
x^0_1 = \frac{1}{\exp{\left[ 1/\sigma -  \ln\chi  \left(\frac{1}{2}\right)\right]}  - \ln\chi}
\end{equation}
%\begin{align}
%P\left(g_i>g_c \right) =
%1 - \frac{1}{2}\exp{\left[-1/\sigma + \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }%&\mbox{if } \\
%%1-\frac{1}{2}\exp{\left[-1/\sigma + \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }&\mbox{if } 
%%\end{cases}
%\end{align}
%otherwise.

>>>>

\begin{document}
%\begin{frontmatter}
%\preprint{}
\title{Optimal strategies of influence in structured social groups}
\author{Colin J. Torney}
\affiliation{Centre for Mathematics \& the Environment, University of Exeter, Penryn Campus, Cornwall, UK}
%\affiliation{Department of Ecology \& Evolutionary Biology, Princeton University, \\Princeton, New Jersey 08544}
\author{Simon A. Levin}
\affiliation{Department of Ecology \& Evolutionary Biology, Princeton University, \\Princeton, New Jersey 08544}

@@

Each individual, $i$, holds an opinion $S_i \in \{ -1,1\}$ that is observable by others in its local neighborhood and is correct when $S_i = \mbox{sign}(G)$. Individuals update their states based on social and personal information. The two information sources are combined such that individual $i$ will adopt the opinion $S_i = 1$ iff
\begin{equation}
\frac{g_i}{\sigma} >  \chi  \left(\frac{1}{2}-x_L\right) 
\label{opt_rule}
\end{equation}
where $x_L$ is the fraction of neighbors with the state $S = 1$ in the local (observable) neighborhood of individual $i$, and $\chi$ is a parameter that determines the weighting individuals give to social information.

While we omit the derivation here the rule defined by Eqn.~\ref{opt_rule} is optimal under certain assumptions\cite{nitzan1982optimal,torney2015}, namely that individuals are aware of the noise in their personal estimate, have a constant level of confidence that neighbors are correct, and assume that the opinions of their neighbors are independent. The result of Eqn.~\ref{opt_rule} is that the probability of an individual switching to the state $S_i = 1$ is a sigmoidal function of the fraction of local neighbors in that state, with the steepness of the transition determined by the values of $\chi$ and $\sigma$ (see Fig~\ref{figSwitch} for a plot of this function).

plot of function

show escape time depends on network structure and clustering

escape time can be predicted if rho is known at the minima

optimal strategy of influence if there's a cost for reducing sigma

and a Bayesian decision rule \cite{nitzan1982optimal,perez2011collective, Perreault, torney2015} is employed. This means that an individual will switch to the state that is the most likely to be correct, given all available information. Formally, individual $i$ will adopt the state $S_i=1$ if
\begin{equation}
 P\left(G \ge 0\big| g_i, \mathbf{S}_{j\in \mathcal{N}_i}\right)  > P\left(G < 0\big| g_i, \mathbf{S}_{j\in \mathcal{N}_i}\right)
\label{opt_sol1}
\end{equation}
where $\mathcal{N}_i$ defines the social interaction neighborhood. To impose a decision rule as a function of $g_i$ we rearrange Eqn.~\ref{opt_sol1} using Bayes' rule, assume individuals place equal weighting on the opinions of neighbors, that they know the level of noise in the environmental cue, and that it is symmetric around zero. Eqn.~\ref{opt_sol1} may then be written as,
\begin{equation}
\frac{\exp{ \left( - \frac{\vert g_i -  \infty \vert}{\sigma} \right)}}{\exp{ \left( - \frac{\vert g_i + \infty \vert}{\sigma} \right)}} > \chi^{1-2x_L}.
\label{rule}
\end{equation}
Here we have denoted the fraction of neighbors an individual observes in the state $S=1$ as $x_L$, used $\pm \infty$ as estimates of the actual value of $G$ (these will cancel), and introduced a social weighting variable $\chi$ which may be interpreted as the belief that a single neighbor is correct rescaled by the neighborhood size. This rescaling means decisions are based on the fraction of neighbors that are correct in the local neighborhood and not the absolute number, ensuring a comparison can be made between social networks with varying average degree.

Rearranging Eqn.~\ref{rule} leads to a decision rule such that

@@

Numerical calculations of $x_2-x_1$ and the optimal strategy are shown in Fig.~\ref{figx2x1}. More intuition into the underlying processes that control the distance between the inflexion points may be attained by performing an expansion of the social information, $\mathcal{F}(x)$, in terms of the moments of the sampling process. Taylor expanding the inner term %$\exp{\left[ - \ln\chi  \left(\frac{1}{2}-x_L\right)\right]}$ 
of Eqn.~\ref{social} around $x_L=x$ leads to % allows us to replace the summation over all values of $x_L$ with its mean and variance, i.e.
\begin{equation}
\mathcal{F}(x) = \frac{1}{2}\exp{\left[ - \ln\chi  \left(\frac{1}{2}-x\right)\right]} \left( 1 + \frac{ \ln(\chi)^2}{2}\mbox{Var}(x_L) \right)
\end{equation}
Effectively this equation approximates the process of observing a restricted subset of the population. The observations of opinions of neighbors represents an estimate, $x_L$, of the true population average opinion, $x$. This random process of observations may be approximated in terms of the mean and variance of the sampling. While the underlying network structure is complex and non-Markovian, it may be modelled as correlated binomial, in essence we model the process as sampling with pairwise correlation, when there is transitivity near-neighbors are likely to have copied one another hence opinions will be clustered. Following ~\cite{witt2004moody}, the variance of a correlated binomial is
\begin{equation}
\mbox{Var}(x_L) = x(1-x)\left( \rho + \frac{1-\rho}{N_s} \right)
\end{equation}
where $\rho$ is the pairwise correlation coefficient. Assuming the network structure and conformity create a positive correlation of opinions amongst neighbors, then we may define an increased likelihood that neighbor $j$ is in state $S_j=1$, given $S_k=1$, so that
\begin{equation}
P(S_j=1|S_k=1) = x(1+\alpha)
\end{equation}
where $\alpha$ is determined by the transitivity (probability $k$ is in the neighborhood of $j$) and conformity. This means 
\begin{equation}
\rho = \frac{x\alpha}{1-x}
\end{equation}
and
\begin{align}
\mathcal{F}(x) &= \frac{1}{2}\exp{\left[ -  \ln\chi  \left(\frac{1}{2}-x\right)\right]} \times \nonumber \\ & \left( 1 + \frac{ \ln(\chi)^2}{2}\left(\frac{x(1-x(1+\alpha))}{N_s} +  \alpha x^2  \right)   \right).
\end{align}
Solutions to $\mathcal{F}(x)=x$ may be constructed with a perturbation method, with small parameters $\alpha$ and $1/N_s$. We begin with 
\begin{equation}
x_1 \sim x^0_1 + \alpha x^1_1 + 1/N_s x^2_1
\end{equation}
The zeroth order solution is for the mean field where there is no clustering and no noise in the sampling process, hence
\begin{align}
x^0  = \frac{1}{2}\exp{\left[ 1/\sigma -  \ln\chi  \left(\frac{1}{2}-x^0\right)\right]} 
\end{align}
For small x
\begin{equation}
x^0_1 = \frac{1}{2\sqrt{\chi}(1-1/\sigma) -  \ln\chi}
\end{equation}
For x close to 0.5 we make use of the fact that $\ln(2x) \sim 2(x-\frac{1}{2})$
\begin{equation}
x^0_2 = \frac{1+1/\sigma - \ln\chi/2 }{2 -  \ln\chi}
\end{equation}

%\begin{align}
%P\left(g_i>g_c \right) =
%1 - \frac{1}{2}\exp{\left[-1/\sigma + \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }%&\mbox{if } \\
%%1-\frac{1}{2}\exp{\left[-1/\sigma + \ln\chi  \left(\frac{1}{2}-x_L\right)\right] }&\mbox{if } 
%%\end{cases}
%\end{align}
%otherwise.
